---
title: "Hauck-Donner"
author: "Göran Broström"
date: '2015-06-06'
output: ioslides_presentation
---

## The Hauck-Donner effect, data

```{r hdata, fig.height=4}
dat = data.frame(T = 1:10, x = c(1:4, 5.02, 4.98, 7:10), event = rep(1, 10))
dat = dat[order(dat$x), ]
with(dat, plot(x, T, type = "b", col = "blue", lty = 2))
```

## Linear regression, code

```{r code, results='hide'}
options(show.signif.stars = FALSE)
fit <- lm(T ~ x, data = dat)
summary(fit)
```

## Linear regression, results

```{r linres, echo=FALSE}
summary(fit)
```

## LR test in linear regression

```{r lrlin}
drop1(fit, test = "F")
```

*Note* that the **F test** is appropriate for _linear models_ (with unknown variance).

*Note* further that the **F** and **Wald** tests are equivalent!

*Warning* for *missing values* (with drop1)!

## ANOVA

```{r anova}
anova(fit)
```

## Cox regression

```{r coxreg, message=FALSE}
library(eha)
fit = coxreg(Surv(T, event) ~ x, data = dat)
fit
```

## Cox regression, drop1

```{r drop1an}
drop1(fit, test = "Chisq")
```

 *Note* the **Hauck-Donner** effect! 
 
 Strikes in *non-linear* models (not only Cox)!